Mouse Development and Aging 
File 3
Gene Onotology (GO) analysis
Last updated Oct 15 2024


```{r}
library(Seurat)
library(SeuratObject) 
library(tidyverse)
library(patchwork)
library(dplyr)
library(tidyr)
library(ggplot2)

#The following packages are all installed via BiocManager. 
#Use the following code (editing as appropriate) to install any missing packages.
#if (!require("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")
#BiocManager::install("clusterProfiler")

library(scRNAtoolVis) #Volcano Plots; requires tidyverse
library(clusterProfiler) #enrichGO
library(org.Mm.eg.db) #gene ontology
library(enrichplot) #gene ontology visualization
library(EnrichedHeatmap) #gene ontology visualization
library(readr) #required for rrvgo
library(rrvgo) #gene ontology visualization

marker <- readRDS("all_degs_p21.rds")


```
###Step 1 - Volcano Plot
Initial visualization both to visualize the DEGs and also to verify data has loaded correctly.

```{r}
#colnames(marker)[colnames(marker) == "Cluster"] <- "cluster"
#jjVolcano(diffData = marker)
#This runs but is messy. Let's try the other volcano. 

#let's rename the clusters for clarity
marker <- marker %>%
  mutate(cluster = ifelse(cluster == "Cortex_adult_vs_P21", "Cortex", cluster))
marker <- marker %>%
  mutate(cluster = ifelse(cluster == "Hippocampus_adult_vs_P21", "Hippocampus", cluster))
marker <- marker %>%
  mutate(cluster = ifelse(cluster == "Thalamus_adult_vs_P21", "Thalamus", cluster))
marker <- marker %>%
  mutate(cluster = ifelse(cluster == "Hypothalamus_adult_vs_P21", "Hypothalamus", cluster))
marker <- marker %>%
  mutate(cluster = ifelse(cluster == "Striatum_adult_vs_P21", "Striatum", cluster))
marker <- marker %>%
  mutate(cluster = ifelse(cluster == "Ventricular System", "Ventricular.Sys", cluster))#renamed this twice so it would display on the graph

#VolcanoPlot
markerVolcano(markers = marker,
              topn = 3,
              labelCol = ggsci::pal_npg()(9))
ggsave('./Figures/DEG/volcano_adult_vs_p21.png', height = 5, width = 8, unit = 'in')
#that's much cleaner


```
###Step 2 - Gene Ontology (1.5)
Next, let us consider gene ontology analysis of the whole brain of our p21 samples compared with those of the adult samples.

We will do this at two different thresholds: 
(1) avg_log2FC = 1.5 
(2) avg_log2FC = 2

Please adjust the code in the first line of code and the name of the saved images when switching from 1.5 to 2. 

```{r}
final_DEGs_1.5 <- marker %>% dplyr::filter(p_val_adj < 0.05, abs(avg_log2FC) > log2(1.5), pct.1 > 0.1, pct.2 > 0.1)

assign("eg", bitr(rownames(final_DEGs_1.5), fromType="SYMBOL", toType="ENTREZID",OrgDb="org.Mm.eg.db"))

      ego <- enrichGO(gene = eg$ENTREZID,
                      OrgDb = org.Mm.eg.db,
                      ont = "BP",
                      pAdjustMethod = "BH",
                      pvalueCutoff  = 0.05,
                      qvalueCutoff = 0.05,
                      readable = TRUE)
      if (nrow(ego@result[ego@result$p.adjust < 0.05 & ego@result$qvalue < 0.05,]) <= 0){
        next
      }
      
      ego <- pairwise_termsim(ego)
      ego <- simplify(ego) 
      
      
     go_analysis <- ego@result[ego@result$p.adjust < 0.05 & ego@result$qvalue < 0.05,]

      simMatrix <- calculateSimMatrix(go_analysis$ID,
                                      orgdb="org.Mm.eg.db",
                                      ont="BP",
                                      method="Rel")

      scores <- setNames(-log10(go_analysis$qvalue), go_analysis$ID)
      reducedTerms <- reduceSimMatrix(simMatrix,
                                      scores,
                                      threshold=0.7,
                                      orgdb="org.Mm.eg.db")


dotplot(ego, showCategory=15)
ggsave('./Figures/GO/GO_dotplot_p21vsadult_1.5.png',   height = 7, width = 7, unit = 'in')

#Generate a barplot sorting by descending order
ego@result <- ego@result %>% arrange(desc(Count))

barplot(ego, showCategory=15)
ggsave('./Figures/GO/GO_barplot_p21vsadult_1.5.png',   height = 7, width = 7, unit = 'in')

#Alternative VolcanoPlot
jjVolcano(diffData = final_DEGs_1.5,
          log2FC.cutoff = log(1.5,2),
          adjustP.cutoff = 0.05,
          topGeneN = 0,
          tile.col = c("#F8766D","#B79F00","#00BA38","#00BFC4",'purple','orange'),
          base_size  = 18,
          legend.position = "none",
          pSize = 1.5,
          celltypeSize = 5
          ) +
   xlab("") +
   ylab("Average log2FoldChange") 

ggsave("./Figures/GO/jjVolcano-1.5FC-p21.png", height = 7, width = 11, unit = 'in')

```


```{r}

CTX <- subset(final_DEGs_1.5, cluster == 'Cortex')
HY <- subset(final_DEGs_1.5, cluster == 'Hypothalamus')
TH <- subset(final_DEGs_1.5, cluster == 'Thalamus')
HP <- subset(final_DEGs_1.5, cluster == 'Hippocampus')
STR <- subset(final_DEGs_1.5, cluster == 'Striatum')
VS <- subset(final_DEGs_1.5, cluster == 'Ventricular.Sys')


#run this block for each region
ego_HY <- enrichGO(
  gene = HY$gene, 
  OrgDb = org.Mm.eg.db, 
  keyType = "SYMBOL", 
  ont = "BP", 
  pAdjustMethod = "BH", 
  pvalueCutoff = 0.05, 
  qvalueCutoff = 0.05, 
  readable = TRUE
)

write.csv(ego@result, file = "./HY_1.5FC.GO.BP.csv")
 

files <- c("./GO/CTX_1.5FC.GO.BP.csv","./GO/HY_1.5FC.GO.BP.csv","./GO/TH_1.5FC.GO.BP.csv","./GO/HP_1.5FC.GO.BP.csv","./GO/STR_1.5FC.GO.BP.csv","./GO/VS_1.5FC.GO.BP.csv")


# Read files into a list of data frames
CTXgo <- read.csv("./GO/CTX_1.5FC.GO.BP.csv")
#CTXgo <- CTXgo[,c("ID","pvalue")]
#write.csv(CTXgo,'./GO/go_terms_CTX.csv',row.names = F)
HYgo <- read.csv("./GO/HY_1.5FC.GO.BP.csv")
#HYgo <- HYgo[,c("ID","pvalue")]
#write.csv(HYgo,'./GO/go_terms_HY.csv',row.names = F)
THgo <- read.csv("./GO/TH_1.5FC.GO.BP.csv")
#THgo <- THgo[,c("ID","pvalue")]
#write.csv(THgo,'./GO/go_terms_TG.csv',row.names = F)
HPgo <- read.csv("./GO/HP_1.5FC.GO.BP.csv")
#HPgo <- HPgo[,c("ID","pvalue")]
#write.csv(HPgo,'./GO/go_terms_HP.csv',row.names = F)
STRgo <- read.csv("./GO/STR_1.5FC.GO.BP.csv")
#STRgo <- STRgo[,c("ID","pvalue")]
#write.csv(STRgo,'./GO/go_terms_STR.csv',row.names = F)
VSgo <- read.csv("./GO/VS_1.5FC.GO.BP.csv")
#VSgo <- VSgo[,c("ID","pvalue")]
#write.csv(VSgo,'./GO/go_terms_VS.csv',row.names = F)

CTXgo$Cluster <- 'CTX'
HYgo$Cluster <- 'HY'
THgo$Cluster <- 'TH'
HPgo$Cluster <- 'HP'
STRgo$Cluster <- 'STR'
VSgo$Cluster <- 'VS'


# Combine all data frames into one
combined_ego_results <- bind_rows(CTXgo, HYgo, THgo, HPgo, STRgo, VSgo)

# Select top 5 terms for each cluster based on adjusted p-value
top_terms <- combined_ego_results %>%
  group_by(Cluster) %>%
  top_n(-5, wt = p.adjust)


ggplot(top_terms, aes(x = Description, y = -log10(p.adjust), color = Cluster)) +
  geom_point(aes(size = Count)) +
  coord_flip() +
  labs(x = "", y = "-log10(Adjusted p-value)", title = "P21 VS. Adult DEG GO Terms by Cluster") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 16),
    axis.text.y = element_text(size = 16),
    axis.title = element_text(size = 16),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 16),
    plot.title = element_text(size = 16, hjust = 0.5)
  )
ggsave("./Figures/GO/dot_1.5FC-p21.png", bg = "white", height = 7, width = 7, unit = 'in')

#try again
ggplot(top_terms, aes(x = Description, y = -log10(p.adjust), color = Cluster)) +
  geom_point(aes(size = Count)) +
  coord_flip() +
  labs(x = "", y = "-log10(Adjusted p-value)", title = "TDP vs AD DEG GO Terms by Cluster") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 16),
    axis.text.y = element_text(size = 16),
    axis.title = element_text(size = 16),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 16),
    plot.title = element_text(size = 16, hjust = 0.5)
  )


```