##Mouse Development and Aging 
###File 1
###Loading through clustering steps

###Introduction
We will compare the development of mouse hippocampal cells at different ages of mice. These samples were collected as controls for various other projects. 

FA: 21-days (n=2) ("young" mice)
CFC: 12-15 weeks (n=2) ("adult" mice)


```{r}
#Load packages used in other projects; some are currently unused but are included as they may be useful in the future. 

library(Seurat)
library(SeuratObject) 
library(hdf5r) 
library(rlang)
library(ggplot2)
library(harmony)
library(RColorBrewer)
library(tidyverse)
library(patchwork)
library(dplyr)
library(tidyr)
library(gridExtra)
library(ggvenn)
library(ComplexUpset)
library(clustree)
library(imager) #save.image
```


###Step 1a - Load Adult Mice Data
First, we will load the data for the adult mice. There were 2 individuals used as controls, so we will load each individually and then merge the datasets.

```{r}
adult_1 <- Load10X_Spatial(
  data.dir = "C:/Users/Spouse/Documents/Research/Xie/CFC/Spatial/home_cage_rep1/outs/",
  slice = "adult_1",
)
adult_1$orig.ident <- "adult_1"
#The above is Yu's FA code. It did not fix the image pt-size display issue. However, all 5 clusters are now appearing in the key for the adult slices, and the adult slices are named correctly.

adult_2 <- Load10X_Spatial(
  data.dir = "C:/Users/Spouse/Documents/Research/Xie/CFC/Spatial/home_cage_rep2/outs/",
  slice = "adult_2",
)
adult_2$orig.ident <- "adult_2"

merge.obj <- merge(adult_1,c(adult_2),
                   add.cell.ids = c("adult_1","adult_2"))
```                  

###Step 1b - Load Young Mice Data & Merge All Mice Data
Now we will merge the data for the adult mice with the data for the young mice. The young mice data was shared via a Seurat object of the two merged repetitions.

```{r}
load("~/Research/Xie/High Folate/Control.Rdata")
merge.obj <- merge(merge.obj,c(controls),
                   add.cell.ids = c("adult","young"))
merge.obj <- JoinLayers(merge.obj)
#Liduo shared the above line of code to ensure 4 images displayed after clustering

#remove unnecessary objects to free up memory
rm(adult_1, adult_2, controls)
```

##Step 2a - Summary Statistics & Quality Control Check
Let's check the genes per spot and UMIs per spot to verify the integrity of the data. Let's also examine the distribution of genes per spot and UMIs per spot across reads to ensure the quality of the dataset.

```{r}
summary(merge.obj$nFeature_Spatial)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#    220    4955    5838    5848    6758   10180 

summary(merge.obj$nCount_Spatial)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#    237   11612   15946   18046   22085   80922 

#The standard is to remove spots with less than 250 genes per spot, as these are low quality reads. Let's see how many spots are low quality reads in this dataset. 

sum(merge.obj$nFeature_Spatial <= 250) #1 - Excellent!

#Likewise, it is standard to remove spots with less than 500 UMIs per spot. Let's see how frequently this occurred. 

sum(merge.obj$nCount_Spatial <= 500) #4 - Great again!

####Let's also visualize the distribution of genes per spot and UMIs per spot with violin plots of the merged dataset, and then individual source datasets.

# metadata <- merge.obj@meta.data

# p1 <- ggplot(metadata, aes(x = "Genes per Spot", y = nFeature_Spatial)) +
#   geom_violin(fill = "maroon", color = "black", alpha = 0.5) +
#   geom_jitter(alpha = 0.2) +
#   ggtitle("Genes per Spot in Merged Data") +
#   theme_minimal() +
#   theme(legend.position = "none", 
#         plot.title = element_text(hjust = 0.5), #Center the title
#         axis.title.x = element_blank(), #Remove x-axis title
#         axis.title.y = element_blank(), #Remove x-axis title
#         axis.text.x = element_blank())  #Remove x-axis text
# 
# p2 <- ggplot(metadata, aes(x = "UMIs per Spot in Merged Data", y = nCount_Spatial)) +
#   geom_violin(fill = "orange", color = "black") +
#   geom_jitter(alpha = 0.2) +
#   ggtitle("UMIs per Spot in Merged Data") +
#   theme_minimal() +
#   theme(legend.position = "none", 
#         plot.title = element_text(hjust = 0.5), #Center the title
#         axis.title.x = element_blank(), #Remove x-axis title
#         axis.title.y = element_blank(), #Remove x-axis title
#         axis.text.x = element_blank())  #Remove x-axis text
#Depricated by VlnPlots below.


####Next, let's look for 3 marker genes to ensure consistency across samples.
color_palette <- c("mf_p21_rep1" = "red", "mf_p21_rep2" = "yellow", "adult_1" = "blue", "adult_2" = "green")
#This ensures the samples are coded with the same colors consistently across our visualizations.

VlnPlot(merge.obj, features = c("nFeature_Spatial"), group.by = "orig.ident", pt.size = 0)+
  scale_fill_manual(values = color_palette)+ guides(fill = FALSE) #guides - show/hide legend
ggsave('./Figures/Genes_Violin/Vln.nFeature.tiff', device='tiff', dpi=500, 
       height = 5, width = 5, unit = 'in')

VlnPlot(merge.obj, features = c("nCount_Spatial"), group.by = "orig.ident",pt.size = 0)+
  scale_fill_manual(values = color_palette) + guides(fill = FALSE)
ggsave('./Figures/Genes_Violin/Vln.nCount.tiff', device='tiff', dpi=500, 
       height = 5, width = 5, unit = 'in')
#These look good! Let's proceed.
```

###Step 2b - Remove Low Quality Spots & Normalize

Let's remove the lowest quality spots (i.e. <=250 genes or <=500 UMIs per spot), normalize via counts per million, select the 3000 most frequent genes and scale the expression values of genes.
```{r}
#Remove the lowest quality spots (Genes <=250 or UMIs <= 500)
merge.obj <- subset(merge.obj, subset = nFeature_Spatial > 250 & nCount_Spatial > 500)

#Normalize to gene frequency in Counts per Million (CPM)
merge.obj <- NormalizeData(merge.obj, assay = "Spatial", scale.factor = 10**6) 

#Next, let's reduce our dataset to focus on only the 3000 genes of highest relative abundance.
merge.obj <- FindVariableFeatures(merge.obj, assay = "Spatial", nfeatures = 3000) 
#Scale the expression values of genes 
merge.obj <- ScaleData(merge.obj, assay = "Spatial", features = rownames(merge.obj))
```
###Step 3a - Noise Reduction - PCA
Let's select the 15 most common PCs and graph an elbow plot to look for a breakpoint. 

```{r}
merge.obj <- RunPCA(merge.obj, assay = "Spatial", npcs = 15, features = rownames(merge.obj))

#ElbowPlot(merge.obj,ndims = 20) + 
#  geom_hline(yintercept = 4.75) + #This may be our breakpoint
#  scale_y_continuous(breaks = c(3,5,10,15,20))
```
###Step 3b - Noise Reduction - Harmony
Using visual analysis of the above, 10, 12, 15, or 19 could be used as breakpoints. Clustering results were poor with 10 PCs, so let's proceed with 15 PCs. 

Next, let's perform harmony to reduce the noise from joining separate datasets.

```{r}
merge.obj <- RunHarmony(merge.obj, "orig.ident", dims.use = 1:15, assay.use = "Spatial") 
#plotHarmony <- ElbowPlot(merge.obj, ndims = 10, reduction = "harmony") + 
#  geom_hline(yintercept = 3) +
#  scale_y_continuous(breaks = c(3, 5, 10, 15))
#plotHarmony
```
###Step 4a - Clustering - UMAP
The 6th component of Harmony was less than the 7th component of Harmony. This is possible but unusual. Harmony is not essential for UMAP, so we examined UMAP with and without Harmony. The results were poor without Harmony, therefore we proceded with Harmony. 

```{r}
#With Harmony
merge.obj <- RunUMAP(merge.obj, dims = 1:15, reduction = "harmony")
#plot - overlay all 4 controls on a single plot
UMAP_harmony <- DimPlot(merge.obj, group.by = "orig.ident")



```
###Step 4b - Clustering - Clustree

```{r}

merge.obj <- FindNeighbors(merge.obj, reduction = "harmony", dims = 1:15)
merge.obj <- FindClusters(merge.obj, resolution = c(0.05, 0.075, 0.093, 0.1, 0.15, 0.2, 0.25, 0.3),  algorithm = 3)
clustree(merge.obj)
ggsave("clustree_15pcs_20240924.tiff", device='tiff', dpi=500, height = 5, width = 5, unit = 'in')
#With 15 PCA's, we are getting 5 clusters at 0.075 and 7 clusters at 0.10. The goal is 6 clusters. 
#It appears 0.093 is where we first find 6 clusters.




```
Other analyses within our lab of mouse brains have used 6 clusters. This corresponds to 0.093. 

```{r}
#The following code was used to compare the SLM method & the Louvian method
#SpatialDimPlot of the SLM method
#slm_dim_0.1 <- SpatialDimPlot(merge.obj, group.by = "Spatial_snn_res.0.1", ncol = 4,label = F, image.alpha = 1, alpha = 1, pt.size.factor = 740)
#pt size 740 for p21
#pt size 3 for adult

#SpatialDimPlot of Louvian method
#dim_0.1 <- SpatialDimPlot(merge.obj, group.by = "Spatial_snn_res.0.1", ncol = 2,label = F, image.alpha = 1, alpha = 1, pt.size.factor = 1)
#ggsave("SpatialRes0.1.png")
#dim_0.1

dim_0.093a <- SpatialDimPlot(merge.obj, group.by = "Spatial_snn_res.0.093", ncol = 2,label = F, image.alpha = 1, alpha = 1, pt.size.factor = 3)
ggsave("dim_0.093a_15pcs_0924.png")

dim_0.093y <- SpatialDimPlot(merge.obj, group.by = "Spatial_snn_res.0.093", ncol = 2,label = F, image.alpha = 1, alpha = 1, pt.size.factor = 740)
ggsave("dim_0.093y_15pcs_0924.png")


dim_0.1a <- SpatialDimPlot(merge.obj, group.by = "Spatial_snn_res.0.1", ncol = 2,label = F, image.alpha = 1, alpha = 1, pt.size.factor = 3)
ggsave("dim_0.1a_15pcs_0924.png")

dim_0.1y <- SpatialDimPlot(merge.obj, group.by = "Spatial_snn_res.0.1", ncol = 2,label = F, image.alpha = 1, alpha = 1, pt.size.factor = 740)
ggsave("dim_0.1y_15pcs_0924.png")


dim_0.15a <- SpatialDimPlot(merge.obj, group.by = "Spatial_snn_res.0.15", ncol = 2,label = F, image.alpha = 1, alpha = 1, pt.size.factor = 3)
ggsave("dim_0.15a_15pcs_0924.png")

dim_0.15y <- SpatialDimPlot(merge.obj, group.by = "Spatial_snn_res.0.15", ncol = 2,label = F, image.alpha = 1, alpha = 1, pt.size.factor = 740)
ggsave("dim_0.15y_15pcs_0924.png")


UMAP_cluster <- DimPlot(merge.obj, group.by = "seurat_clusters")

#save(merge.obj, file = "MD_merge.obj.Rdata")
#saves current output to speed up future work
```


```{r}

```